### 2015-12-17 document/manuscript-plos.tex
PLoS formatted manuscript

% Template for PLoS
% Version 3.1 February 2015

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% fixltx2e package for \textsubscript
\usepackage{fixltx2e}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% rotating package for sideways tables
\usepackage{rotating}

% Remove comment for double spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\sf PLOS}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

%% END MACROS SECTION

\begin{document}
\vspace*{0.35in}

% Title must be 250 characters or less.
% Please capitalize all terms in the title except conjunctions, prepositions, and articles.
\begin{flushleft}
{\Large
\textbf\newline{Ten Simple Rules for Taking Advantage of GitHub in Bioinformatics}
}
\newline
\\
Yasset Perez-Riverol\textsuperscript{1,*},
Rui Wang\textsuperscript{1},
Timo Sachsenberg\textsuperscript{2},
Julian Uszkoreit\textsuperscript{3},
Laurent Gatto\textsuperscript{4},
Felipe da Veiga Leprevost\textsuperscript{5},
Christian Fufezan\textsuperscript{6},
Juan Antonio Vizcaíno\textsuperscript{1,*}

\bigskip

\bf{1} European Molecular Biology Laboratory, European Bioinformatics
Institute (EMBL-EBI), Wellcome Trust Genome Campus, Hinxton,
Cambridge, CB10 1SD, UK.
\\
\bf{2} Applied Bioinformatics and Department of Computer Science,
University of Tübingen, D-72074 Tübingen, Germany.
\\
\bf{3} Medizinisches Proteom-Center, Ruhr-Universität Bochum,
Universitätsstr. 150, D-44801 Bochum, Germany.
\\
\bf{4} Computational Proteomics Unit, Cambridge Systems Biology
Centre, University of Cambridge Tennis Court Road Cambridge, CB2 1GA,
UK.
\\
\bf{5} Department of Pathology, University of Michigan, Ann Arbor,
Michigan 48109, USA.
\\
\bf{6} Institute of Plant Biology and Biotechnology, University of
Muenster, Schlossplatz 8, 48143 Muenster, Germany.
\bigskip

% Use the asterisk to denote corresponding authorship and provide email address in note below.
* yperez@ebi.ac.uk and juan@ebi.ac.uk

\end{flushleft}

% Please keep the abstract below 300 words
% \section*{Abstract}
% Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur
% eget porta erat. Morbi consectetur est vel gravida
% pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam
% sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus
% mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc
% pellentesque pretium diam, quis maximus dolor faucibus id. Nunc
% convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet
% enim ultrices, ultrices elit pulvinar, volutpat risus.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
% \section*{Author Summary}
% Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur
% eget porta erat. Morbi consectetur est vel gravida
% pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam
% sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus
% mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc
% pellentesque pretium diam, quis maximus dolor faucibus id. Nunc
% convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet
% enim ultrices, ultrices elit pulvinar, volutpat risus.

\linenumbers

\subsection*{Introduction}\label{introduction}

Bioinformatics is a broad discipline in which the common denominator is
the need to produce and/or use software that can be applied to
biological data in different contexts. To enable and ensure the
reproducibility of scientific claims, it is essential that the
scientific publication, the corresponding datasets and the data analysis
are made publicly available \cite{Goodman:2014,Perez-Riverol:2015}. All
software used for the analysis should be either carefully documented
(e.g.~for commercial software) or, better, openly shared and directly
accessible to others \cite{Osborne:2014,Vihinen:2015}. The rise of
openly available software and source code, and concomitant collaborative
development is facilitated by the existence of several code repository
services such as SourceForge (http://sourceforge.net/), Bitbucket
(https://bitbucket.org/) and GitHub (https://github.com/), among others.
These resources are also essential for collaborative software projects,
since they enable the organisation and sharing of programming tasks
between different remote contributors. Here, we introduce the main
features of GitHub, a popular web-based platform which offers a free and
integrated environment for hosting the source code, documentation and
web page for open source projects. GitHub also offers paid plans for
private repositories. GitHub relies, at its core, on the well-known and
open source version control system git, designed and developed by Linus
Torvalds for the development of the Linux kernel. One reason for
GitHub's success is that it offers more than a simple source code
hosting service. It provides developers with a dynamic and collaborative
environment, often coined as social coding platform, with the ability to
review, comment and discuss code \cite{Dabbish:2012}.

Some of our recommendations outlined below are applicable to other
hosting services. However our main aim here is to highlight specific
GitHub features. We provide a set of recommendations to take full
advantage of GitHub's features to manage small and large bioinformatics
projects and increase their profile and visibility. These rules have
been ordered to reflect a typical development process: learning git and
GitHub basics, use branches, label and tag code, track the project using
issues, etc.

\subsection*{Rule 1. Structure your projects: users, organisations,
repositories and
teams}\label{rule-1.-structure-your-projects-users-organisations-repositories-and-teams}

Open source projects on GitHub are visible to everyone, but write
permissions, i.e.~the possibility to directly modify the content, need
to explicitly be granted. On the other hand, everyone with a GitHub
account can fork any public project and start developing in one's own
fork. This forking is the basis of social coding. It allows anyone to
develop and test novel features into existing code and offers the
possibility to merge novel features back the into the main project,
thereby becoming a contributor. Structuring your projects allows to
manage permissions and restrict access at different levels: users, teams
and organisations. Users are the keystone of GitHub, as for any other
social network. Every user has a profile listing their GitHub projects
and activities, which can be populated optionally with personal
information including name, e-mail address, image and webpage. To stay
up to date with the activity of other users one can \emph{follow} their
accounts. Collaboration can be achieved by simply adding a trusted
\emph{Collaborator} and thereby granting write access. However,
development in large projects is usually done by teams of people, within
a larger organisation. GitHub organisations are a great way to manage
team-based access permissions for the individual projects of institutes,
research labs, and large open source projects that need multiple owners
and administrators (Fig. 1). We recommend that you (as an individual
researcher) make your profile visible to other users and display all the
projects and organisations you are working in, including a list of the
latest activities on the site (Fig. 1). Finally, repositories (the
shortened term is \emph{repo}) are versioned directories or dedicated
storage spaces for your software projects, which can be included inside
an organisation or can belong to particular users. Users can usually
keep code, text files, images and small data files inside a repo. And
while many users store programs and code projects, there is nothing
preventing users from keeping text documents such as analysis reports
and manuscripts (see for example the repository for this manuscipt at
https://github.com/ypriverol/github-paper), or other file types in your
projects. Note that until recently, GitHub was lacking support for
storing large files (\textgreater{}100 MB), a issue that has been
recently addressed by the GitHub large file storage.

\begin{figure}[htbp]
\centering
% \includegraphics[width=\linewidth]{../figures/figure01_overview.pdf}
\caption{The structure of a GitHub-based project illustrating project
structure and interactions with the community.}
\end{figure}

\subsection*{Rule 2. Learn Git and embrace its
power}\label{rule-2.-learn-git-and-embrace-its-power}

The cornerstone of GitHub is the distributed version control system git.
Every change, from fixing a typo to a complete redesign of the software
is controlled by versions, so called revisions. While beginners may
consider the learning curve of Git steep, many introductory and detailed
tutorials are available. A revision can be considered as a
\emph{snapshot} (version) of a file system. Git is remarkably effective
in archiving the complete history of a project (all revisions) by,
amongst other things, storing only the differences among them. To create
a new revision, the set of changes introduced (e.g.~new, deleted or
modified files) are committed to the repository. Following the rule:
``commit often, as most as you can, perfection later'', one can keep
track of the development in small incremental changes. At any time it is
possible to go back to a previous version. In larger projects, multiple
users contribute to the same repository.

\subsection*{Rule 3. Use branches}\label{rule-3.-use-branches}

Concurrent development including commits to the same repository can be
organised using several common approaches. The most common way is to use
git \emph{branches} to separate different lines of development. Active
development is often performed on a development branch and stable
versions e.g.~those used for a software release, are kept in a master
branch. In practice, developers work on one or several features or
improvements. To keep commits of the different features logically
separated, distinct branches are typically used. Later, when development
is complete and none of the tests fail (see Rule 6) new features can be
merged back into the development line or master branch. During such a
development, the original branch might continuously be developed and
other features might be merged into the master branch. Nevertheless, one
can always pull the currently up-to date master branch into one fork,
always enabling to react or adapt to the changes in the code. In
projects involving more than one contributor, everyone wants to be sure
that the contributions of others increase the quality and move the
project forward. \emph{Forking} a repository and providing \emph{pull
requests} constitute an easy way for collaboration inside and over
organisations boundaries. A user that forks a repository creates a copy
under their GitHub account. Modifications like a branch with new
features or bug fixes can conveniently be provided to the forked
(upstream) repository by opening a pull request. Once it is opened for
review and discussion, it usually results in additional insights and in
an increased code quality. Once a pull request gets accepted, typically
it gets merged into the development branch.

\subsection*{Rule 4. Use tags and semantic version
numbering}\label{rule-4.-use-tags-and-semantic-version-numbering}

Tags offer the possibility to label versions during the development
process, tracking its progress. Version numbering should follow semantic
versioning in the form X.Y.Z, with X being the major, Y the minor and Z
the patch version of the release, including possible meta information,
as described in http://semver.org/. Correct labelling allows developers
and users to easily recover older versions, compare them, or simply use
them to reproduce results described in publications (see Rule 8). This
approach will also help defining a coherent software publication
strategy (see Rule 6).

\subsection*{Rule 5. The code must always be ready to use: continuously
integrate}\label{rule-5.-the-code-must-always-be-ready-to-use-continuously-integrate}

The first rule of software development is that the code needs to be
ready to use as soon as possible \cite{Leprevost:2014}, remain so during
development, and should be well-documented and tested. In 2005, Martin
Fowler defined the basic principles for continuous integration in
software development \cite{FowlerCI}. These principles have become the
main reference for best practices in continuous integration, providing
the framework needed to deploy software, and in some way also data.
Every repository, script, mathematical model, and function should
contain a set of self-automated tests. A source code may run, but that
does not mean it is doing the right thing. The simple use of those
self-automated tests is to detect possible bugs introduced by new
features, or changes in the code or dependencies, but also to detect
wrong results, the so called \emph{logic errors}, where the source code
produces a different result compared to what one intended it to do.
Then, continuous integration provides the way of automatically run all
of these tests in the repository by checking data and software
dependencies. Continuous integration can be done automatically on GitHub
(See Rule 6).

\subsection*{Rule 6: Automate for better code
quality}\label{rule-6-automate-for-better-code-quality}

GitHub offers different types of hooks that are executed after each push
to a repository, making easier to follow the basic principles of
continuous integration. The GitHub web hooks allows third-party
platforms to access and interact with a GitHub repository and thus to
automate post-processing tasks. Such tasks demonstrate to the community
that a project follows rigorous software engineering processes, often
associated with high quality development. It also shows that is
currently working and has documentation that reflects the current code.
We suggest that all these three tasks become part of your project.
Firstly, continuous integration can be achieved by \emph{Travis}
(https://travis-ci.org), a hosted continued integration platform that is
free for all open source projects. Travis builds and tests the source
code using a plethora of options such as different platforms and
interpreter versions. Furthermore it offers notifications which allow
your team and contributors to know if the new changes work, and prevent
the introduction of errors in the code, making the repo always ready to
use. Secondly, in addition to successful completion of the tests, one
can also demonstrate that they cover the existing code base
sufficiently. For this task, the integration of \emph{Codecov} is
recommended (https://codecov.io). Thirdly, one might consider to
automatically update the documentation upon code modification. This
implies that your projects provide comprehensive documentation so others
can understand, and contribute back to them. For Python or C/C++ code,
automatic documentation generation can be done using sphinx
(http://sphinx-doc.org/) and subsequently integrated into GitHub using
``Read the Docs'' (https://readthedocs.org/). All of these platforms
will create reports and badges (also called shields) for the projects
that can be included on your GitHub page, thereby making your projects
easily identifiable as high quality and well-maintained.

\subsection*{Rule 7. Use and maintain your GitHub issue
trackers}\label{rule-7.-use-and-maintain-your-github-issue-trackers}

GitHub \emph{issues} are a great way to keep track of bugs, tasks, and
enhancements. Classical issue trackers are primarily intended to be used
as bug trackers. In contrast, GitHub issues follow a different
philosophy: each tracker has its own section in every repository, and
can be used to trace bugs, new ideas, and enhancements, by using a
powerful but optional tagging system for each issue. Its main focus is
put in promoting collaboration, providing context by using
cross-references, and an excellent text formatting for each issue: (i) a
title and description, (ii) colour-coded labels help to categorise and
filter issues, (iii) milestones act as a container for issues,(iv) one
assignee responsible for working on the issue, and (v) comments that
allow anyone with access to the repository to provide feedback. Another
aspect is its simplicity. For instance, it does not require to fill
lengthy forms including every piece of information that might be
valuable to reproduce the bug. It only requires to give the title and
provide some optional text. If the developer needs more information,
they can simply request it in a comment. GitHub issues are then more
dynamic and pose a lower barrier for users to report bugs and request
features. A well-organised and tagged issue tracker will help upcoming
contributors and users to understand a project more deeply.

\subsection*{Rule 8. Make your code easily citable, and cite source
code!}\label{rule-8.-make-your-code-easily-citable-and-cite-source-code}

In research, it is a good practice to ensure permanent and unambiguous
identifiers for citable items like articles, datasets, or biological
entities such as proteins, genes and metabolites, among others. Digital
Object Identifiers (DOIs) have been used for many years as unique and
unambiguous identifiers for enabling the citation of scientific
publications. More recently, a trend has started to produce DOIs for
other types of scientific outputs such as datasets \cite{Vizcaino:2014}
or training materials (for example \cite{Ahmadia_2015_27353}). The main
motivation behind this is to give scientists a better credit for their
work \cite{NatBiotechEditorial:2009}, enabling at the same time a better
way to cite and track it. A common issue with software is that it
normally evolves at a different speed than what is published in the
scientific literature. In fact, it is common to find software having
novel features and functionalities that were not described in the
original publication. GitHub now enables the use of DOIs to cite the
code deposited, using the data archiving tool Zenodo
(https://zenodo.org/). The procedure is simple (see
https://guides.GitHub.com/activities/citable-code/) and, by default,
Zenodo takes an archive of a repository each time a new release is
created in GitHub. However, before Zenodo can issue a DOI, metadata
needs to be provided about the archived repository. Once the DOI has
been assigned, apart from using it in your CV, you can add it to
literature information resources such as Europe PubMed Central
\cite{EuropePMCConsortium:2015}. As already mentioned in the
introduction, reproducibility of scientific claims should be enabled by
providing openly the software, the datasets and the process leading to
interpretable results that are used in a particular study. One should
always highlight as much as possible in publications that the code is
freely available in, for example, GitHub, together with any other
relevant piece of information that may have been deposited. In our
experience, this openness substantially increases your chances of
getting the paper accepted for publication. On one hand, journal editors
and reviewers have the opportunity to reproduce your findings during the
manuscript review process, increasing the confidence of your results. On
the other hand, once the paper is published, your work can be reproduced
by any member of the scientific community, which can increase citations
and foster opportunities for further discussion and collaboration. Also
one must have in mind that the availability of a public repository with
the source code does not make the software open source \emph{per se}, as
it needs to have an appropriate license
(http://opensource.org/licenses/).

\subsection*{Rule 9. Promote your projects in the scientific community -
create a web page and
more}\label{rule-9.-promote-your-projects-in-the-scientific-community---create-a-web-page-and-more}

Additional steps can be done to boost the visibility of a organisation.
For example, GitHub \emph{Pages} are simple landing webpages that GitHub
hosts for free without the need for a server or database. GitHub users
can create and host blog websites, help pages, manuals, tutorials and
websites related to specific projects. \emph{Pages} comes with a
powerful static site generator called Jekyll (https://jekyllrb.com) that
can be integrated with other platforms such as Bootstrap
(http://getbootstrap.com/) or Disqus (https://disqus.com/), to support
and moderate comments. In addition, GitHub also provides mechanisms for
real-time communication called Gitter (http://gitter.im). Gitter is a
GitHub-based chat tool (in limited beta at the time of writing) which
enables developers and users to share aspects of their work. Gitter
inherits the shape of the social groups operating around GitHub
repositories, organisations, and issues. It relies on the identity
within GitHub, creating IRC (Internet Relay Chat)-like chat rooms for
public and private repositories. From within a Gitter chat, members can
reference issues, comments, or pull-requests. A different service is
Gist (https://gist.github.com), which represents a unique way to share
\emph{code snippets}, single files, parts of files, or full
applications. Gist can be generated in two different ways: public
\emph{gists}, that can be browsed and searched, and secret gists that
are not provided through \emph{Discover}
(https://gist.github.com/discover). One of the main features of Gist is
the possibility to embed code snippets in other applications, enabling
users to embed gists in any text field that supports JavaScript.

\subsection*{Rule 10. Check periodically existing open source
projects}\label{rule-10.-check-periodically-existing-open-source-projects}

One of the main tasks of scientists is to actively follow the
developments in their field. Analogously, scientific programmers need to
revise publicly available (e.g.~open source) projects and code that can
be interesting for their research. Therefore, you should try to learn as
much as possible from your peers and keep up-to-date with all the
developments of relevant projects. GitHub enables this functionality by
\emph{following} other GitHub users (also mentioned in Rule 1) or
\emph{watching} the activity of projects, which is a common feature in
many social media platforms. Take advantage of it as much as possible!

\subsection*{Conclusions}\label{conclusions}

If you are interested and have not used GitHub so far, we recommend you
to get started as soon as possible. As in any other topic, a learning
curve is required for beginners. However, we anticipate the reward will
be worth your effort.

\subsection*{Disclaimer}\label{disclaimer}

The authors have no affiliation with GitHub, nor any commercial entity
mentioned in this article. The views described here reflect our own
without any input from any third party organisation.

\nolinenumbers

\bibliography{refs.bib}

\end{document}

### 2016-03-31 document/manuscript-plos.tex
lots of wording changes

@@
 Bioinformatics is a broad discipline in which the common denominator is
-the need to produce and/or use software that can be applied to
+the need to produce and\slash or use software that can be applied to
 biological data in different contexts. To enable and ensure the
@@
 software used for the analysis should be either carefully documented
-(e.g.~for commercial software) or, better, openly shared and directly
+(e.g., for commercial software) or, better, openly shared and directly
 accessible to others \cite{Osborne:2014,Vihinen:2015}. The rise of
@@
 These resources are also essential for collaborative software projects,
-since they enable the organisation and sharing of programming tasks
+since they enable the organization and sharing of programming tasks
 between different remote contributors. Here, we introduce the main
-features of GitHub, a popular web-based platform which offers a free and
-integrated environment for hosting the source code, documentation and
+features of GitHub, a popular web-based platform that offers a free and
+integrated environment for hosting the source code, documentation, and
 web page for open source projects. GitHub also offers paid plans for
-private repositories. GitHub relies, at its core, on the well-known and
+private repositories. At its core, GitHub relies on the well-known and
 open source version control system git, designed and developed by Linus
@@
 hosting service. It provides developers with a dynamic and collaborative
-environment, often coined as social coding platform, with the ability to
+environment, often referred to as social coding platform, with the ability to
 review, comment and discuss code \cite{Dabbish:2012}.
@@
 advantage of GitHub's features to manage small and large bioinformatics
-projects and increase their profile and visibility. These rules have
-been ordered to reflect a typical development process: learning git and
-GitHub basics, use branches, label and tag code, track the project using
+projects and increase their profile and visibility. These rules are
+ordered to reflect a typical development process: learning git and
+GitHub basics, using branches, labeling and tagging code, tracking the project using
 issues, etc.
 
-\subsection*{Rule 1. Structure your projects: users, organisations,
+\subsection*{Rule 1. Structure your projects: users, organizations,
 repositories and
@@
 
-Open source projects on GitHub are visible to everyone, but write
+Public repositories for open source projects on GitHub are visible to everyone, but write
 permissions, i.e.~the possibility to directly modify the content, need
-to explicitly be granted. On the other hand, everyone with a GitHub
-account can fork any public project and start developing in one's own
+to be granted explicitly. On the other hand, anyone with a GitHub
+account can fork any public project and start developing in their own
 fork. This forking is the basis of social coding. It allows anyone to
-develop and test novel features into existing code and offers the
-possibility to merge novel features back the into the main project,
-thereby becoming a contributor. Structuring your projects allows to
+develop and test novel features in existing code and offers the
+possibility of merging that work back the into the main project,
+thereby becoming a contributor. Structuring your projects allows you to
 manage permissions and restrict access at different levels: users, teams
-and organisations. Users are the keystone of GitHub, as for any other
+and organizations. Users are the central element of GitHub, as in any other
 social network. Every user has a profile listing their GitHub projects
 and activities, which can be populated optionally with personal
-information including name, e-mail address, image and webpage. To stay
-up to date with the activity of other users one can \emph{follow} their
+information including name, e-mail address, image, and webpage. To stay
+up to date with the activity of other users, one can \emph{follow} their
 accounts. Collaboration can be achieved by simply adding a trusted
@@
 development in large projects is usually done by teams of people, within
-a larger organisation. GitHub organisations are a great way to manage
+a larger organization. GitHub organizations are a great way to manage
 team-based access permissions for the individual projects of institutes,
@@
 researcher) make your profile visible to other users and display all the
-projects and organisations you are working in, including a list of the
+projects and organizations you are working in, including a list of the
 latest activities on the site (Fig. 1). Finally, repositories (the
-shortened term is \emph{repo}) are versioned directories or dedicated
+short term is \emph{repo}) are versioned directories or dedicated
 storage spaces for your software projects, which can be included inside
-an organisation or can belong to particular users. Users can usually
-keep code, text files, images and small data files inside a repo. And
+an organization or can belong to particular users. Users can usually
+keep code, text files, images, and small data files inside a repo. And
 while many users store programs and code projects, there is nothing
 preventing users from keeping text documents such as analysis reports
-and manuscripts (see for example the repository for this manuscipt at
-https://github.com/ypriverol/github-paper), or other file types in your
-projects. Note that until recently, GitHub was lacking support for
-storing large files (\textgreater{}100 MB), a issue that has been
+and manuscripts (see for example the repository for this manuscript at
+\url{https://github.com/ypriverol/github-paper}), or other file types in your
+projects. Note that until recently, GitHub lacked support for
+storing large files (\textgreater{}100 MB), but this has been
 recently addressed by the GitHub large file storage.
@@
 \centering
-% \includegraphics[width=\linewidth]{../figures/figure01_overview.pdf}
+% \includegraphics[width=\linewidth]{./figure01_overview.pdf}
 \caption{The structure of a GitHub-based project illustrating project
@@
 
-The cornerstone of GitHub is the distributed version control system git.
-Every change, from fixing a typo to a complete redesign of the software
-is controlled by versions, so called revisions. While beginners may
-consider the learning curve of Git steep, many introductory and detailed
-tutorials are available. A revision can be considered as a
-\emph{snapshot} (version) of a file system. Git is remarkably effective
-in archiving the complete history of a project (all revisions) by,
-amongst other things, storing only the differences among them. To create
-a new revision, the set of changes introduced (e.g.~new, deleted or
+The backbone of GitHub is the distributed version control system git.
+Every change, from fixing a typo to a complete redesign of the software,
+is stored as a version, called a revision. While beginners may
+find the learning curve for git steep, many introductory and detailed
+tutorials are available. A revision can be viewed as a
+\emph{snapshot} (version) of a file system. Git is remarkably efficient
+at archiving the complete history of a project (all revisions) by,
+among other things, storing only the differences among them. To create
+a new revision, the set of changes introduced (e.g., new, deleted, or
 modified files) are committed to the repository. Following the rule:
-``commit often, as most as you can, perfection later'', one can keep
+``commit often, as much as you can, perfect later'', one can keep
 track of the development in small incremental changes. At any time it is
@@
 
-Concurrent development including commits to the same repository can be
-organised using several common approaches. The most common way is to use
+Concurrent development, including multiple commits to the same repository, can be
+organized using several common approaches. The most common way is to use
 git \emph{branches} to separate different lines of development. Active
-development is often performed on a development branch and stable
-versions e.g.~those used for a software release, are kept in a master
-branch. In practice, developers work on one or several features or
+development is often performed on a development branch, and stable
+versions, e.g., those used for a software release, are kept in a master
+branch. In practice, developers often work concurrently on one or several features or
 improvements. To keep commits of the different features logically
 separated, distinct branches are typically used. Later, when development
-is complete and none of the tests fail (see Rule 6) new features can be
+is complete and none of the tests fail (see Rule 6), new features can be
 merged back into the development line or master branch. During such a
@@
 can always pull the currently up-to date master branch into one fork,
-always enabling to react or adapt to the changes in the code. In
+always enabling one to react or adapt to the changes in the code. In
 projects involving more than one contributor, everyone wants to be sure
 that the contributions of others increase the quality and move the
-project forward. \emph{Forking} a repository and providing \emph{pull
-requests} constitute an easy way for collaboration inside and over
-organisations boundaries. A user that forks a repository creates a copy
+project forward. \emph{Forking} a repository and using \emph{pull
+requests} constitutes an easy method for collaboration inside and over
+organizational boundaries. A user who forks a repository creates a copy
 under their GitHub account. Modifications like a branch with new
@@
 review and discussion, it usually results in additional insights and in
-an increased code quality. Once a pull request gets accepted, typically
-it gets merged into the development branch.
+increased code quality. Once a pull request is accepted, it typically
+is merged into the development branch.
 
@@
 
-Tags offer the possibility to label versions during the development
-process, tracking its progress. Version numbering should follow semantic
-versioning in the form X.Y.Z, with X being the major, Y the minor and Z
+Tags offer the possibility of labeling versions during the development
+process, to track its progress. Version numbering should follow semantic
+versioning in the form X.Y.Z, with X being the major, Y the minor, and Z
 the patch version of the release, including possible meta information,
-as described in http://semver.org/. Correct labelling allows developers
+as described in \url{http://semver.org/}. Correct labelling allows developers
 and users to easily recover older versions, compare them, or simply use
 them to reproduce results described in publications (see Rule 8). This
-approach will also help defining a coherent software publication
+approach also helps to define a coherent software publication
 strategy (see Rule 6).
@@
 The first rule of software development is that the code needs to be
-ready to use as soon as possible \cite{Leprevost:2014}, remain so during
-development, and should be well-documented and tested. In 2005, Martin
+ready to use as soon as possible \cite{Leprevost:2014}, to remain so during
+development, and to be well-documented and tested. In 2005, Martin
 Fowler defined the basic principles for continuous integration in
@@
 features, or changes in the code or dependencies, but also to detect
-wrong results, the so called \emph{logic errors}, where the source code
-produces a different result compared to what one intended it to do.
-Then, continuous integration provides the way of automatically run all
+wrong results, often known as \emph{logic errors}, where the source code
+produces a different result than what was intended.
+Continuous integration provides a way to automatically run all
 of these tests in the repository by checking data and software
@@
 GitHub offers different types of hooks that are executed after each push
-to a repository, making easier to follow the basic principles of
-continuous integration. The GitHub web hooks allows third-party
+to a repository, making it easier to follow the basic principles of
+continuous integration. The GitHub web hooks allow third-party
 platforms to access and interact with a GitHub repository and thus to
 automate post-processing tasks. Such tasks demonstrate to the community
-that a project follows rigorous software engineering processes, often
-associated with high quality development. It also shows that is
+that a project follows rigorous software engineering processes, which are often
+associated with high quality development. They also show that the project
 currently working and has documentation that reflects the current code.
-We suggest that all these three tasks become part of your project.
-Firstly, continuous integration can be achieved by \emph{Travis}
+We suggest that all three of the following tasks become part of your project.
+First, continuous integration can be achieved by \emph{Travis}
 (https://travis-ci.org), a hosted continued integration platform that is
@@
 code using a plethora of options such as different platforms and
-interpreter versions. Furthermore it offers notifications which allow
+compiler\slash interpreter versions. In addition, it offers notifications that allow
 your team and contributors to know if the new changes work, and prevent
 the introduction of errors in the code, making the repo always ready to
-use. Secondly, in addition to successful completion of the tests, one
-can also demonstrate that they cover the existing code base
-sufficiently. For this task, the integration of \emph{Codecov} is
-recommended (https://codecov.io). Thirdly, one might consider to
-automatically update the documentation upon code modification. This
-implies that your projects provide comprehensive documentation so others
-can understand, and contribute back to them. For Python or C/C++ code,
+use. Second, in addition to successfully completing the tests, one
+can also demonstrate that the tests sufficiently cover the existing code base.
+For this, integration of \emph{Codecov} (\url{https://codecov.io}) is
+recommended. Third, one might consider
+automatically updating the documentation upon code modification. This
+implies that your project provides comprehensive documentation so that others
+can understand, and contribute back to it. For Python or C/C++ code,
 automatic documentation generation can be done using sphinx
-(http://sphinx-doc.org/) and subsequently integrated into GitHub using
-``Read the Docs'' (https://readthedocs.org/). All of these platforms
-will create reports and badges (also called shields) for the projects
+(\url{http://sphinx-doc.org/}) and subsequently integrated into GitHub using
+``Read the Docs'' (\url{https://readthedocs.org/}). These tools
+create reports and badges (also called shields) for the projects
 that can be included on your GitHub page, thereby making your projects
@@
 powerful but optional tagging system for each issue. Its main focus is
-put in promoting collaboration, providing context by using
-cross-references, and an excellent text formatting for each issue: (i) a
-title and description, (ii) colour-coded labels help to categorise and
-filter issues, (iii) milestones act as a container for issues,(iv) one
+in promoting collaboration, providing context by using
+cross-references.  Issues also have very clear formatting: (i) a
+title and description, (ii) color-coded labels help to categorize and
+filter issues, (iii) milestones act as a container for issues, (iv) an
 assignee responsible for working on the issue, and (v) comments that
 allow anyone with access to the repository to provide feedback. Another
-aspect is its simplicity. For instance, it does not require to fill
-lengthy forms including every piece of information that might be
-valuable to reproduce the bug. It only requires to give the title and
+aspect of issues is its simplicity. For instance, it does not require someone to fill
+out lengthy forms that include every piece of information that might be
+valuable to reproduce the bug. It only requires someone to give the title and
 provide some optional text. If the developer needs more information,
-they can simply request it in a comment. GitHub issues are then more
-dynamic and pose a lower barrier for users to report bugs and request
-features. A well-organised and tagged issue tracker will help upcoming
+they can simply request it in a comment. GitHub issues are thus more
+dynamic and pose a lower barrier-to-entry for users to report bugs and request
+features. A well-organized and tagged issue tracker will help upcoming
 contributors and users to understand a project more deeply.
@@
 or training materials (for example \cite{Ahmadia_2015_27353}). The main
-motivation behind this is to give scientists a better credit for their
-work \cite{NatBiotechEditorial:2009}, enabling at the same time a better
+motivation behind this is to give scientists better credit for their
+work \cite{NatBiotechEditorial:2009}, simultaneously enabling a better
 way to cite and track it. A common issue with software is that it
-normally evolves at a different speed than what is published in the
+normally evolves at a different speed than text published in the
 scientific literature. In fact, it is common to find software having
@@
 code deposited, using the data archiving tool Zenodo
-(https://zenodo.org/). The procedure is simple (see
-https://guides.GitHub.com/activities/citable-code/) and, by default,
-Zenodo takes an archive of a repository each time a new release is
+(\url{https://zenodo.org/}). The procedure is simple (see
+\url{https://guides.GitHub.com/activities/citable-code/}) and, by default,
+Zenodo creates an archive of a repository each time a new release is
 created in GitHub. However, before Zenodo can issue a DOI, metadata
@@
 introduction, reproducibility of scientific claims should be enabled by
-providing openly the software, the datasets and the process leading to
-interpretable results that are used in a particular study. One should
-always highlight as much as possible in publications that the code is
-freely available in, for example, GitHub, together with any other
-relevant piece of information that may have been deposited. In our
+openly providing the software, the datasets, and the process leading to
+interpretable results that was used in a particular study. As much as possible,
+publications should highlight that the code is
+freely available in, for example, GitHub, and cite the DOI of the archived version
+of the code that was used, together with any other
+relevant information that may have been deposited. In our
 experience, this openness substantially increases your chances of
@@
 and reviewers have the opportunity to reproduce your findings during the
-manuscript review process, increasing the confidence of your results. On
+manuscript review process, increasing their confidence in your results. On
 the other hand, once the paper is published, your work can be reproduced
@@
 and foster opportunities for further discussion and collaboration. Also
-one must have in mind that the availability of a public repository with
+one must keep in mind that the availability of a public repository with
 the source code does not make the software open source \emph{per se}, as
-it needs to have an appropriate license
-(http://opensource.org/licenses/).
+it also needs to have an appropriate license
+(\url{http://opensource.org/licenses/}).
 
@@
 
-Additional steps can be done to boost the visibility of a organisation.
+Additional steps can boost the visibility of a organization.
 For example, GitHub \emph{Pages} are simple landing webpages that GitHub
-hosts for free without the need for a server or database. GitHub users
-can create and host blog websites, help pages, manuals, tutorials and
+hosts for free without the need for a separate server or database. GitHub users
+can create and host blog websites, help pages, manuals, tutorials, and
 websites related to specific projects. \emph{Pages} comes with a
-powerful static site generator called Jekyll (https://jekyllrb.com) that
+powerful static site generator called Jekyll (\url{https://jekyllrb.com}) that
 can be integrated with other platforms such as Bootstrap
-(http://getbootstrap.com/) or Disqus (https://disqus.com/), to support
-and moderate comments. In addition, GitHub also provides mechanisms for
-real-time communication called Gitter (http://gitter.im). Gitter is a
-GitHub-based chat tool (in limited beta at the time of writing) which
+(\url{http://getbootstrap.com/}) or Disqus (\url{https://disqus.com/}), to support
+and moderate comments. In addition, GitHub also provides a mechanism for
+real-time communication called Gitter (\url{http://gitter.im}). Gitter is a
+GitHub-based chat tool (in limited beta at the time of writing) that
 enables developers and users to share aspects of their work. Gitter
-inherits the shape of the social groups operating around GitHub
-repositories, organisations, and issues. It relies on the identity
+inherits the network of social groups operating around GitHub
+repositories, organizations, and issues. It relies on the identity
 within GitHub, creating IRC (Internet Relay Chat)-like chat rooms for
-public and private repositories. From within a Gitter chat, members can
-reference issues, comments, or pull-requests. A different service is
-Gist (https://gist.github.com), which represents a unique way to share
+public and private repositories. Within a Gitter chat, members can
+reference issues, comments, or pull requests. A different service is
+Gist (\url{https://gist.github.com}), which represents a unique way to share
 \emph{code snippets}, single files, parts of files, or full
-applications. Gist can be generated in two different ways: public
-\emph{gists}, that can be browsed and searched, and secret gists that
+applications. Gists can be generated in two different ways: public
+\emph{gists} that can be browsed and searched, and secret gists that
 are not provided through \emph{Discover}
-(https://gist.github.com/discover). One of the main features of Gist is
-the possibility to embed code snippets in other applications, enabling
+(\url{https://gist.github.com/discover}). One of the main features of Gist is
+the ability to embed code snippets in other applications, enabling
 users to embed gists in any text field that supports JavaScript.
@@
 
-One of the main tasks of scientists is to actively follow the
+One of scientists' main tasks is to actively follow the
 developments in their field. Analogously, scientific programmers need to
-revise publicly available (e.g.~open source) projects and code that can
+revise publicly available (e.g., open source) projects and code that might
 be interesting for their research. Therefore, you should try to learn as
-much as possible from your peers and keep up-to-date with all the
+much as possible from your peers, and keep up-to-date with all the
 developments of relevant projects. GitHub enables this functionality by
@@
 If you are interested and have not used GitHub so far, we recommend you
-to get started as soon as possible. As in any other topic, a learning
-curve is required for beginners. However, we anticipate the reward will
+get started as soon as possible. As with any other topic, there is a learning
+curve for beginners. However, we anticipate the reward will
 be worth your effort.
@@
 mentioned in this article. The views described here reflect our own
-without any input from any third party organisation.
+without any input from any third party organization.
 

### 2016-03-31 document/manuscript-plos.tex
adding a ref for software citation

@@
 publications. More recently, a trend has started to produce DOIs for
-other types of scientific outputs such as datasets \cite{Vizcaino:2014}
-or training materials (for example \cite{Ahmadia_2015_27353}). The main
+other types of scientific outputs such as datasets \cite{Vizcaino:2014},
+training materials (for example \cite{Ahmadia_2015_27353}), and software. The main
 motivation behind this is to give scientists better credit for their
-work \cite{NatBiotechEditorial:2009}, simultaneously enabling a better
+work \cite{NatBiotechEditorial:2009,force11-scwg}, simultaneously enabling a better
 way to cite and track it. A common issue with software is that it

### 2016-04-01 document/manuscript-plos.tex

--- 2015-12-17 document/manuscript-plos.tex
@@
 \usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
+
+\usepackage{longtable}
+\usepackage{booktabs}
+\usepackage{array}
 
@@
 {\Large
-\textbf\newline{Ten Simple Rules for Taking Advantage of GitHub in Bioinformatics}
+\textbf\newline{Ten Simple Rules for Taking Advantage of git and GitHub}
 }
@@
 \\
-Yasset Perez-Riverol\textsuperscript{1,*},
+
+Yasset Perez-RiverolYasset Perez-Riverol\textsuperscript{1,*}
+Laurent Gatto\textsuperscript{2},
 Rui Wang\textsuperscript{1}, 
-Timo Sachsenberg\textsuperscript{2},
-Julian Uszkoreit\textsuperscript{3},
-Laurent Gatto\textsuperscript{4},
+Timo Sachsenberg \textsuperscript{3},
+Julian Uszkoreit\textsuperscript{4},
 Felipe da Veiga Leprevost\textsuperscript{5},
 Christian Fufezan\textsuperscript{6},
+Tobias Ternent\textsuperscript{1},
+Stephen J. Eglen\textsuperscript{7}, 
+Daniel S. Katz\textsuperscript{8}, 
+Tom Pollard\textsuperscript{9}, 
+Alexander Konovalov\textsuperscript{10},
 Juan Antonio Vizcaíno\textsuperscript{1,*}
@@
 \\
-\bf{2} Applied Bioinformatics and Department of Computer Science,
+\bf{2} Computational Proteomics Unit, Cambridge Systems Biology
+Centre, University of Cambridge, Tennis Court Road, Cambridge, CB2
+1GA, UK.
+\\
+\bf{3} Applied Bioinformatics and Department of Computer Science,
 University of Tübingen, D-72074 Tübingen, Germany.
 \\
-\bf{3} Medizinisches Proteom-Center, Ruhr-Universität Bochum,
+\bf{4} Medizinisches Proteom-Center, Ruhr-Universität Bochum,
 Universitätsstr. 150, D-44801 Bochum, Germany.
-\\
-\bf{4} Computational Proteomics Unit, Cambridge Systems Biology
-Centre, University of Cambridge Tennis Court Road Cambridge, CB2 1GA,
-UK.
 \\
@@
 Muenster, Schlossplatz 8, 48143 Muenster, Germany.
+\\
+\bf{7} Centre for Mathematical Sciences, University of Cambridge,
+Wilberforce Road, Cambridge CB3 0WA, UK.
+\\
+\bf{8} National Center for Supercomputing Applications, University of
+Illinois, 1205 W. Clark St., Urbana, Illinois 61801, USA.
+\\
+\bf{9} MIT Laboratory for Computational Physiology, Institute for
+Medical Engineering and Science, Massachusetts Institute of
+Technology, E25-545B, 45 Carleton St, Cambridge, MA 02142, USA.
+\\
+\bf{10} Centre for Interdisciplinary Research in Computational
+Algebra, University of St Andrews, St Andrews, Fife, KY16 9SX, UK.
 \bigskip
@@
 
-\subsection*{Introduction}\label{introduction}
-
-Bioinformatics is a broad discipline in which the common denominator is
-the need to produce and/or use software that can be applied to
-biological data in different contexts. To enable and ensure the
-reproducibility of scientific claims, it is essential that the
-scientific publication, the corresponding datasets and the data analysis
-are made publicly available \cite{Goodman:2014,Perez-Riverol:2015}. All
-software used for the analysis should be either carefully documented
-(e.g.~for commercial software) or, better, openly shared and directly
-accessible to others \cite{Osborne:2014,Vihinen:2015}. The rise of
-openly available software and source code, and concomitant collaborative
-development is facilitated by the existence of several code repository
-services such as SourceForge (http://sourceforge.net/), Bitbucket
-(https://bitbucket.org/) and GitHub (https://github.com/), among others.
-These resources are also essential for collaborative software projects,
-since they enable the organisation and sharing of programming tasks
-between different remote contributors. Here, we introduce the main
-features of GitHub, a popular web-based platform which offers a free and
-integrated environment for hosting the source code, documentation and
-web page for open source projects. GitHub also offers paid plans for
-private repositories. GitHub relies, at its core, on the well-known and
-open source version control system git, designed and developed by Linus
-Torvalds for the development of the Linux kernel. One reason for
-GitHub's success is that it offers more than a simple source code
-hosting service. It provides developers with a dynamic and collaborative
-environment, often coined as social coding platform, with the ability to
-review, comment and discuss code \cite{Dabbish:2012}.
-
-Some of our recommendations outlined below are applicable to other
-hosting services. However our main aim here is to highlight specific
-GitHub features. We provide a set of recommendations to take full
-advantage of GitHub's features to manage small and large bioinformatics
-projects and increase their profile and visibility. These rules have
-been ordered to reflect a typical development process: learning git and
-GitHub basics, use branches, label and tag code, track the project using
-issues, etc.
-
-\subsection*{Rule 1. Structure your projects: users, organisations,
-repositories and
-teams}\label{rule-1.-structure-your-projects-users-organisations-repositories-and-teams}
-
-Open source projects on GitHub are visible to everyone, but write
-permissions, i.e.~the possibility to directly modify the content, need
-to explicitly be granted. On the other hand, everyone with a GitHub
-account can fork any public project and start developing in one's own
-fork. This forking is the basis of social coding. It allows anyone to
-develop and test novel features into existing code and offers the
-possibility to merge novel features back the into the main project,
-thereby becoming a contributor. Structuring your projects allows to
-manage permissions and restrict access at different levels: users, teams
-and organisations. Users are the keystone of GitHub, as for any other
-social network. Every user has a profile listing their GitHub projects
-and activities, which can be populated optionally with personal
-information including name, e-mail address, image and webpage. To stay
-up to date with the activity of other users one can \emph{follow} their
-accounts. Collaboration can be achieved by simply adding a trusted
-\emph{Collaborator} and thereby granting write access. However,
-development in large projects is usually done by teams of people, within
-a larger organisation. GitHub organisations are a great way to manage
-team-based access permissions for the individual projects of institutes,
-research labs, and large open source projects that need multiple owners
-and administrators (Fig. 1). We recommend that you (as an individual
-researcher) make your profile visible to other users and display all the
-projects and organisations you are working in, including a list of the
-latest activities on the site (Fig. 1). Finally, repositories (the
-shortened term is \emph{repo}) are versioned directories or dedicated
-storage spaces for your software projects, which can be included inside
-an organisation or can belong to particular users. Users can usually
-keep code, text files, images and small data files inside a repo. And
-while many users store programs and code projects, there is nothing
-preventing users from keeping text documents such as analysis reports
-and manuscripts (see for example the repository for this manuscipt at
-https://github.com/ypriverol/github-paper), or other file types in your
-projects. Note that until recently, GitHub was lacking support for
-storing large files (\textgreater{}100 MB), a issue that has been
-recently addressed by the GitHub large file storage.
-
-\begin{figure}[htbp]
-\centering
-% \includegraphics[width=\linewidth]{../figures/figure01_overview.pdf}
-\caption{The structure of a GitHub-based project illustrating project
-structure and interactions with the community.}
-\end{figure}
-
-\subsection*{Rule 2. Learn Git and embrace its
-power}\label{rule-2.-learn-git-and-embrace-its-power}
-
-The cornerstone of GitHub is the distributed version control system git.
-Every change, from fixing a typo to a complete redesign of the software
-is controlled by versions, so called revisions. While beginners may
-consider the learning curve of Git steep, many introductory and detailed
-tutorials are available. A revision can be considered as a
-\emph{snapshot} (version) of a file system. Git is remarkably effective
-in archiving the complete history of a project (all revisions) by,
-amongst other things, storing only the differences among them. To create
-a new revision, the set of changes introduced (e.g.~new, deleted or
-modified files) are committed to the repository. Following the rule:
-``commit often, as most as you can, perfection later'', one can keep
-track of the development in small incremental changes. At any time it is
-possible to go back to a previous version. In larger projects, multiple
-users contribute to the same repository.
-
-\subsection*{Rule 3. Use branches}\label{rule-3.-use-branches}
-
-Concurrent development including commits to the same repository can be
-organised using several common approaches. The most common way is to use
-git \emph{branches} to separate different lines of development. Active
-development is often performed on a development branch and stable
-versions e.g.~those used for a software release, are kept in a master
-branch. In practice, developers work on one or several features or
-improvements. To keep commits of the different features logically
-separated, distinct branches are typically used. Later, when development
-is complete and none of the tests fail (see Rule 6) new features can be
-merged back into the development line or master branch. During such a
-development, the original branch might continuously be developed and
-other features might be merged into the master branch. Nevertheless, one
-can always pull the currently up-to date master branch into one fork,
-always enabling to react or adapt to the changes in the code. In
-projects involving more than one contributor, everyone wants to be sure
-that the contributions of others increase the quality and move the
-project forward. \emph{Forking} a repository and providing \emph{pull
-requests} constitute an easy way for collaboration inside and over
-organisations boundaries. A user that forks a repository creates a copy
-under their GitHub account. Modifications like a branch with new
-features or bug fixes can conveniently be provided to the forked
-(upstream) repository by opening a pull request. Once it is opened for
-review and discussion, it usually results in additional insights and in
-an increased code quality. Once a pull request gets accepted, typically
-it gets merged into the development branch.
-
-\subsection*{Rule 4. Use tags and semantic version
-numbering}\label{rule-4.-use-tags-and-semantic-version-numbering}
-
-Tags offer the possibility to label versions during the development
-process, tracking its progress. Version numbering should follow semantic
-versioning in the form X.Y.Z, with X being the major, Y the minor and Z
-the patch version of the release, including possible meta information,
-as described in http://semver.org/. Correct labelling allows developers
-and users to easily recover older versions, compare them, or simply use
-them to reproduce results described in publications (see Rule 8). This
-approach will also help defining a coherent software publication
-strategy (see Rule 6).
-
-\subsection*{Rule 5. The code must always be ready to use: continuously
-integrate}\label{rule-5.-the-code-must-always-be-ready-to-use-continuously-integrate}
-
-The first rule of software development is that the code needs to be
-ready to use as soon as possible \cite{Leprevost:2014}, remain so during
-development, and should be well-documented and tested. In 2005, Martin
-Fowler defined the basic principles for continuous integration in
-software development \cite{FowlerCI}. These principles have become the
-main reference for best practices in continuous integration, providing
-the framework needed to deploy software, and in some way also data.
-Every repository, script, mathematical model, and function should
-contain a set of self-automated tests. A source code may run, but that
-does not mean it is doing the right thing. The simple use of those
-self-automated tests is to detect possible bugs introduced by new
-features, or changes in the code or dependencies, but also to detect
-wrong results, the so called \emph{logic errors}, where the source code
-produces a different result compared to what one intended it to do.
-Then, continuous integration provides the way of automatically run all
-of these tests in the repository by checking data and software
-dependencies. Continuous integration can be done automatically on GitHub
-(See Rule 6).
-
-\subsection*{Rule 6: Automate for better code
-quality}\label{rule-6-automate-for-better-code-quality}
-
-GitHub offers different types of hooks that are executed after each push
-to a repository, making easier to follow the basic principles of
-continuous integration. The GitHub web hooks allows third-party
-platforms to access and interact with a GitHub repository and thus to
-automate post-processing tasks. Such tasks demonstrate to the community
-that a project follows rigorous software engineering processes, often
-associated with high quality development. It also shows that is
-currently working and has documentation that reflects the current code.
-We suggest that all these three tasks become part of your project.
-Firstly, continuous integration can be achieved by \emph{Travis}
-(https://travis-ci.org), a hosted continued integration platform that is
-free for all open source projects. Travis builds and tests the source
-code using a plethora of options such as different platforms and
-interpreter versions. Furthermore it offers notifications which allow
-your team and contributors to know if the new changes work, and prevent
-the introduction of errors in the code, making the repo always ready to
-use. Secondly, in addition to successful completion of the tests, one
-can also demonstrate that they cover the existing code base
-sufficiently. For this task, the integration of \emph{Codecov} is
-recommended (https://codecov.io). Thirdly, one might consider to
-automatically update the documentation upon code modification. This
-implies that your projects provide comprehensive documentation so others
-can understand, and contribute back to them. For Python or C/C++ code,
-automatic documentation generation can be done using sphinx
-(http://sphinx-doc.org/) and subsequently integrated into GitHub using
-``Read the Docs'' (https://readthedocs.org/). All of these platforms
-will create reports and badges (also called shields) for the projects
-that can be included on your GitHub page, thereby making your projects
-easily identifiable as high quality and well-maintained.
-
-\subsection*{Rule 7. Use and maintain your GitHub issue
-trackers}\label{rule-7.-use-and-maintain-your-github-issue-trackers}
-
-GitHub \emph{issues} are a great way to keep track of bugs, tasks, and
-enhancements. Classical issue trackers are primarily intended to be used
-as bug trackers. In contrast, GitHub issues follow a different
-philosophy: each tracker has its own section in every repository, and
-can be used to trace bugs, new ideas, and enhancements, by using a
-powerful but optional tagging system for each issue. Its main focus is
-put in promoting collaboration, providing context by using
-cross-references, and an excellent text formatting for each issue: (i) a
-title and description, (ii) colour-coded labels help to categorise and
-filter issues, (iii) milestones act as a container for issues,(iv) one
-assignee responsible for working on the issue, and (v) comments that
-allow anyone with access to the repository to provide feedback. Another
-aspect is its simplicity. For instance, it does not require to fill
-lengthy forms including every piece of information that might be
-valuable to reproduce the bug. It only requires to give the title and
-provide some optional text. If the developer needs more information,
-they can simply request it in a comment. GitHub issues are then more
-dynamic and pose a lower barrier for users to report bugs and request
-features. A well-organised and tagged issue tracker will help upcoming
-contributors and users to understand a project more deeply.
-
-\subsection*{Rule 8. Make your code easily citable, and cite source
-code!}\label{rule-8.-make-your-code-easily-citable-and-cite-source-code}
-
-In research, it is a good practice to ensure permanent and unambiguous
-identifiers for citable items like articles, datasets, or biological
-entities such as proteins, genes and metabolites, among others. Digital
-Object Identifiers (DOIs) have been used for many years as unique and
-unambiguous identifiers for enabling the citation of scientific
-publications. More recently, a trend has started to produce DOIs for
-other types of scientific outputs such as datasets \cite{Vizcaino:2014}
-or training materials (for example \cite{Ahmadia_2015_27353}). The main
-motivation behind this is to give scientists a better credit for their
-work \cite{NatBiotechEditorial:2009}, enabling at the same time a better
-way to cite and track it. A common issue with software is that it
-normally evolves at a different speed than what is published in the
-scientific literature. In fact, it is common to find software having
-novel features and functionalities that were not described in the
-original publication. GitHub now enables the use of DOIs to cite the
-code deposited, using the data archiving tool Zenodo
-(https://zenodo.org/). The procedure is simple (see
-https://guides.GitHub.com/activities/citable-code/) and, by default,
-Zenodo takes an archive of a repository each time a new release is
-created in GitHub. However, before Zenodo can issue a DOI, metadata
-needs to be provided about the archived repository. Once the DOI has
-been assigned, apart from using it in your CV, you can add it to
-literature information resources such as Europe PubMed Central
-\cite{EuropePMCConsortium:2015}. As already mentioned in the
-introduction, reproducibility of scientific claims should be enabled by
-providing openly the software, the datasets and the process leading to
-interpretable results that are used in a particular study. One should
-always highlight as much as possible in publications that the code is
-freely available in, for example, GitHub, together with any other
-relevant piece of information that may have been deposited. In our
-experience, this openness substantially increases your chances of
-getting the paper accepted for publication. On one hand, journal editors
-and reviewers have the opportunity to reproduce your findings during the
-manuscript review process, increasing the confidence of your results. On
-the other hand, once the paper is published, your work can be reproduced
-by any member of the scientific community, which can increase citations
-and foster opportunities for further discussion and collaboration. Also
-one must have in mind that the availability of a public repository with
-the source code does not make the software open source \emph{per se}, as
-it needs to have an appropriate license
-(http://opensource.org/licenses/).
-
-\subsection*{Rule 9. Promote your projects in the scientific community -
-create a web page and
-more}\label{rule-9.-promote-your-projects-in-the-scientific-community---create-a-web-page-and-more}
-
-Additional steps can be done to boost the visibility of a organisation.
-For example, GitHub \emph{Pages} are simple landing webpages that GitHub
-hosts for free without the need for a server or database. GitHub users
-can create and host blog websites, help pages, manuals, tutorials and
-websites related to specific projects. \emph{Pages} comes with a
-powerful static site generator called Jekyll (https://jekyllrb.com) that
-can be integrated with other platforms such as Bootstrap
-(http://getbootstrap.com/) or Disqus (https://disqus.com/), to support
-and moderate comments. In addition, GitHub also provides mechanisms for
-real-time communication called Gitter (http://gitter.im). Gitter is a
-GitHub-based chat tool (in limited beta at the time of writing) which
-enables developers and users to share aspects of their work. Gitter
-inherits the shape of the social groups operating around GitHub
-repositories, organisations, and issues. It relies on the identity
-within GitHub, creating IRC (Internet Relay Chat)-like chat rooms for
-public and private repositories. From within a Gitter chat, members can
-reference issues, comments, or pull-requests. A different service is
-Gist (https://gist.github.com), which represents a unique way to share
-\emph{code snippets}, single files, parts of files, or full
-applications. Gist can be generated in two different ways: public
-\emph{gists}, that can be browsed and searched, and secret gists that
-are not provided through \emph{Discover}
-(https://gist.github.com/discover). One of the main features of Gist is
-the possibility to embed code snippets in other applications, enabling
-users to embed gists in any text field that supports JavaScript.
-
-\subsection*{Rule 10. Check periodically existing open source
-projects}\label{rule-10.-check-periodically-existing-open-source-projects}
-
-One of the main tasks of scientists is to actively follow the
-developments in their field. Analogously, scientific programmers need to
-revise publicly available (e.g.~open source) projects and code that can
-be interesting for their research. Therefore, you should try to learn as
-much as possible from your peers and keep up-to-date with all the
-developments of relevant projects. GitHub enables this functionality by
-\emph{following} other GitHub users (also mentioned in Rule 1) or
-\emph{watching} the activity of projects, which is a common feature in
-many social media platforms. Take advantage of it as much as possible!
-
-\subsection*{Conclusions}\label{conclusions}
-
-If you are interested and have not used GitHub so far, we recommend you
-to get started as soon as possible. As in any other topic, a learning
-curve is required for beginners. However, we anticipate the reward will
-be worth your effort.
-
-\subsection*{Disclaimer}\label{disclaimer}
-
-The authors have no affiliation with GitHub, nor any commercial entity
-mentioned in this article. The views described here reflect our own
-without any input from any third party organisation.
+\input manuscript
 

### 2016-04-01 document/manuscript-plos.tex
put tables sideways
update to affiliation
Fixed duplicated name

@@
 \usepackage{array}
+\usepackage{rotating}
 
@@
 
-Yasset Perez-RiverolYasset Perez-Riverol\textsuperscript{1,*}
+Yasset Perez-Riverol\textsuperscript{1,*}
 Laurent Gatto\textsuperscript{2},
@@
 \\
-\bf{8} National Center for Supercomputing Applications, University of
+\bf{8} National Center for Supercomputing Applications \&
+Graduate School of Library and Information Science, University of
 Illinois, 1205 W. Clark St., Urbana, Illinois 61801, USA.

### 2016-04-01 document/manuscript-plos.tex

@@
 Alexander Konovalov\textsuperscript{10},
+Robert M Flight\textsuperscript{11},
 Juan Antonio Vizcaíno\textsuperscript{1,*}
@@
 Algebra, University of St Andrews, St Andrews, Fife, KY16 9SX, UK.
+\\
+\bf{11} Department of Molecular Biology and Biochemistry, Markey
+Cancer Center, Resource Center for Stable Isotope-Resolved
+Metabolomics, University of Kentucky, 800 Rose Street Lexington, KY
+40536-0093, USA.
+
 \bigskip

### 2016-04-01 document/manuscript-plos.tex
update name and affiliation

@@
 Daniel S. Katz\textsuperscript{8}, 
-Tom Pollard\textsuperscript{9}, 
+Tom J. Pollard\textsuperscript{9}, 
 Alexander Konovalov\textsuperscript{10},
@@
 Medical Engineering and Science, Massachusetts Institute of
-Technology, E25-545B, 45 Carleton St, Cambridge, MA 02142, USA.
+Technology, Cambridge, MA 02142, USA.
 \\

### 2016-04-01 document/manuscript-plos.tex
move acknowledgements in manusctipt-plos.tex file, document funding sources

@@
 
+\section*{Acknowledgments}
+
+The authors would like to thank C. Titus Brown, Chris Fields and an
+anonymous reviewer for their helpful comments in improving the
+manuscript. We would also like to thank C. Titus Brown from
+advertising contributions to the manuscript on social media, leading
+to further contributions and improvements. We also thank Peter Cock
+(peterjc) and and Kai Blin (kblin) for helpful suggestion contributed
+directly though GitHub.
+
 \nolinenumbers

### 2016-04-02 document/manuscript-plos.tex
add Kai to author list

@@
 Robert M Flight\textsuperscript{11},
+Kai Blin\textsuperscript{12},
 Juan Antonio Vizcaíno\textsuperscript{1,*}
@@
 \\
-\bf{8} National Center for Supercomputing Applications \&
-Graduate School of Library and Information Science, University of
-Illinois, 1205 W. Clark St., Urbana, Illinois 61801, USA.
+\bf{8} National Center for Supercomputing Applications \& Graduate
+School of Library and Information Science, University of Illinois,
+1205 W. Clark St., Urbana, Illinois 61801, USA.
 \\
@@
 40536-0093, USA.
+\\
+\bf{12} The Novo Nordisk Foundation Center for Biosustainability,
+Technical University of Denmark, Kogle Alle 6, 2970 Hørsholm, Denmark
 
@@
 to further contributions and improvements. We also thank Peter Cock
-(peterjc) and and Kai Blin (kblin) for helpful suggestion contributed
-directly though GitHub.
+(peterjc) for helpful suggestion contributed directly though GitHub.
 

### 2016-04-11 document/manuscript-plos.tex
last update/compile before submission
Update manuscript-plos.tex
I have removed the reviewers from the acknowledgments, as requested by PLOS Comp. Biology.

@@
 
-The authors would like to thank C. Titus Brown, Chris Fields and an
-anonymous reviewer for their helpful comments in improving the
-manuscript. We would also like to thank C. Titus Brown from
-advertising contributions to the manuscript on social media, leading
-to further contributions and improvements. We also thank Peter Cock
-(peterjc) for helpful suggestion contributed directly though GitHub.
+The authors would like to thank C. Titus Brown for mentioning
+the manuscript on social media, leading to additional contributions
+and further improvements. We also thank Peter Cock (peterjc) for
+helpful suggestion contributed directly though GitHub.
 

### 2016-05-08 document/manuscript-plos.tex
Update manuscript-plos.tex
I have modified the Affiliations section as requested by PLOS: all affiliations listed contain a Department, Institution, City, and Country (removing information such as street name and post/zip code).

@@
 Institute (EMBL-EBI), Wellcome Trust Genome Campus, Hinxton,
-Cambridge, CB10 1SD, UK.
+Cambridge, UK.
 \\
 \bf{2} Computational Proteomics Unit, Cambridge Systems Biology
-Centre, University of Cambridge, Tennis Court Road, Cambridge, CB2
-1GA, UK.
+Centre, University of Cambridge, Cambridge, UK.
 \\
 \bf{3} Applied Bioinformatics and Department of Computer Science,
-University of Tübingen, D-72074 Tübingen, Germany.
+University of Tübingen, Tübingen, Germany.
 \\
-\bf{4} Medizinisches Proteom-Center, Ruhr-Universität Bochum,
-Universitätsstr. 150, D-44801 Bochum, Germany.
+\bf{4} Medizinisches Proteom-Center, Ruhr-Universität Bochum, Bochum, Germany.
 \\
 \bf{5} Department of Pathology, University of Michigan, Ann Arbor,
-Michigan 48109, USA.
+Michigan, USA.
 \\
 \bf{6} Institute of Plant Biology and Biotechnology, University of
-Muenster, Schlossplatz 8, 48143 Muenster, Germany.
+Muenster, Muenster, Germany.
 \\
 \bf{7} Centre for Mathematical Sciences, University of Cambridge,
-Wilberforce Road, Cambridge CB3 0WA, UK.
+Cambridge, UK.
 \\
@@
 School of Library and Information Science, University of Illinois,
-1205 W. Clark St., Urbana, Illinois 61801, USA.
+Urbana, Illinois, USA.
 \\
@@
 Medical Engineering and Science, Massachusetts Institute of
-Technology, Cambridge, MA 02142, USA.
+Technology, Cambridge, MA, USA.
 \\
 \bf{10} Centre for Interdisciplinary Research in Computational
-Algebra, University of St Andrews, St Andrews, Fife, KY16 9SX, UK.
+Algebra, University of St Andrews, St Andrews, UK.
 \\
@@
 Cancer Center, Resource Center for Stable Isotope-Resolved
-Metabolomics, University of Kentucky, 800 Rose Street Lexington, KY
-40536-0093, USA.
+Metabolomics, University of Kentucky, Lexington, KY, USA.
 \\
 \bf{12} The Novo Nordisk Foundation Center for Biosustainability,
-Technical University of Denmark, Kogle Alle 6, 2970 Hørsholm, Denmark
+Technical University of Denmark, Hørsholm, Denmark.
 
@@
 
+\section*{Supporting Information Legends}
+
+\textbf{Supplementary File S1.} Supplementary Information including
+three sections: Git Large File Storage (LFS), Testing Levels of the
+Source Code and Continuous integration, and Source code documentation.
+
 \end{document}
